A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data

### 2016-11-04 bioc-workflow.Rnw
link to pRolocVis gif

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
  \caption{A screen shot of clickable interface and zoomable PCA plot of the main app in the \Biocpkg{pRolocGUI} package.}
  \label{fig:app}
\end{figure}

As diplayed in the screenshot above, the \textit{main} application is
designed for exploratory data analysis and is divied into 3 tabs: (1)
PCA, (2) Profiles and (3) Table selection. The default view upon
loading is the PCA tab, which features a clickable interface and
zoomable PCA plot with an interactive data table for displaying the
quantitation information. Particular proteins of interest can be
highlighted using the text search box. There is also an alternate
profiles tab for visualisation of the protein profiles, which can be
used to examine the patterns of proteins of interest. The Table
selection tab provides an interface to control data table column
selection.

>>>>

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
  \caption{A screen shot of clickable interface and zoomable PCA plot
    of the main app in the \Biocpkg{pRolocGUI} package. }
  \label{fig:app}
\end{figure}

As diplayed in the screenshot above, the \textit{main} application is
designed for exploratory data analysis and is divied into 3 tabs: (1)
PCA, (2) Profiles and (3) Table selection. The default view upon
loading is the PCA tab, which features a clickable interface and
zoomable PCA plot with an interactive data table for displaying the
quantitation information. Particular proteins of interest can be
highlighted using the text search box. There is also an alternate
profiles tab for visualisation of the protein profiles, which can be
used to examine the patterns of proteins of interest. The Table
selection tab provides an interface to control data table column
selection. A short
animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
illustrating the interface is available in the manuscript repository~\cite{ghrepo}.

### 2016-11-04 bioc-workflow.tex
link to pRolocVis gif

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
## Subset [6725,20][5032,20] Fri Nov  4 11:36:18 2016 
## Removed features with more than 0 NAs: Fri Nov  4 11:36:18 2016 
## Dropped featureData's levels Fri Nov  4 11:36:18 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

% latex table generated in R 3.3.2 by xtable 1.8-2 package
% Fri Nov  4 11:36:20 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127N &   1 & 127N & 1 to 6 (pooled) & 6.00 \\ 
  X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X128N &   1 & 128N & 10 to 11 (pooled) & 13.30 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
  X128N.1 &   2 & 128N & 10 to 11 (pooled) & 12.50 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Fri Nov  4 11:36:19 2016 
## Added markers from  'pdres' marker vector. Fri Nov  4 11:36:20 2016 
##  MSnbase version: 2.0.0
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4344085               0.3070891               0.3823223 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6928348               0.6017963               0.4249301 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4226011               0.5911111               0.9498226 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7957737               0.7096282               0.3152061 
##         Plasma membrane              Proteasome 
##               0.7053008               0.4133848
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      84                     170                      84 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     301                     479                      94 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      27                     123                     523 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     230                     341                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     326                     154                    2058
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
## Subset [6725,20][5032,20] Fri Nov  4 12:03:41 2016 
## Removed features with more than 0 NAs: Fri Nov  4 12:03:41 2016 
## Dropped featureData's levels Fri Nov  4 12:03:41 2016 
##  MSnbase version: 1.21.7
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

% latex table generated in R 3.3.2 by xtable 1.8-2 package
% Fri Nov  4 12:03:43 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127N &   1 & 127N & 1 to 6 (pooled) & 6.00 \\ 
  X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X128N &   1 & 128N & 10 to 11 (pooled) & 13.30 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
  X128N.1 &   2 & 128N & 10 to 11 (pooled) & 12.50 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Fri Nov  4 12:03:43 2016 
## Added markers from  'pdres' marker vector. Fri Nov  4 12:03:43 2016 
##  MSnbase version: 2.0.0
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4369956               0.3026572               0.3845674 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6735594               0.6449643               0.4375399 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4123260               0.6045583               0.9494543 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7957342               0.7090410               0.3143893 
##         Plasma membrane              Proteasome 
##               0.7176337               0.4177169
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      83                     174                      84 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     300                     460                     106 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      27                     121                     524 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     229                     343                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     328                     157                    2058
\end{verbatim}
\end{kframe}
\end{knitrout}
